\documentclass{sig-alternate}

\usepackage{mathtools,MnSymbol}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{cite}
%\usepackage{amsthm}


\allowdisplaybreaks



%\addtolength{\textwidth}{1in}
%\addtolength{\oddsidemargin}{-0.6in}
%\addtolength{\evensidemargin}{-0.6in}
\addtolength{\textheight}{0.30cm}
\addtolength{\topmargin}{-0.15cm}
%\leftmargini 2.9ex


\begin{document}

\input{macros}


% Copyright
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


\title{Jumping the ORDER BY Barrier in Large-Scale Pattern Matching}



%\begin{comment}


\numberofauthors{5} 
\author{
\alignauthor
Daniel Lupei\\
   \affaddr{EPFL}\\
   \email{daniel.lupei@epfl.ch}
\alignauthor
Mike Barnett\\
   \affaddr{MSR, Redmond}\\
   \email{mbarnett@microsoft.com}
\alignauthor
Saeed Maleki\\
   \affaddr{MSR, Redmond}\\
   \email{saemal@microsoft.com}
\and
\alignauthor
Madan Musuvathi\\
   \affaddr{MSR, Redmond}\\
   \email{madanm@microsoft.com}
\alignauthor 
Todd Mytkowicz\\
       \affaddr{MSR, Redmond}\\
       \email{toddm@microsoft.com}
}


%\end{comment}


\maketitle

\begin{abstract}
Event-series pattern matching is a major component of large-scale data 
analytics pipelines enabling a wide range of system diagnostics tasks.  
%While existing solutions focus on the online scenario and expect 
%their data to be (mostly) sorted on time, this is not the case in many 
%situations where batch processing is preferable due to the large volume of data 
%and where the physical data layout has to meet wider system requirements. 
A precursor to pattern matching is an expensive ``shuffle the world'' stage
wherein data are ordered by time and shuffled across the network.  Because many
existing systems treat the pattern matching engine as a black box, they are
unable to optimizing the entire data analytics pipeline, and in particular, this
costly shuffle.

This paper demonstrates how to optimize such queries. We first translate an
expressive class of regular-expression like patterns to relational queries such
that they can benefit from decades of progress in relational optimizers, and
then we introduce the technique of {\em abstract pattern matching}, a linear
time preprocessing step which, adapting ideas from symbolic execution and
abstract interpretation, discards events from the input guaranteed not to appear
in successful matches.  Abstract pattern matching first computes a conservative
representation of the output-relevant domain of every transition in a pattern
based on the (unary) predicates of that transition.  It then further refines
these domains based on the structure of the pattern (i.e., paths through the
pattern) as well as any of the pattern's join predicates across transitions.
The outcome is an {\em abstract filter} that when applied to the original stream
excludes events that are guaranteed not to participate in a match.

We implemented and applied {\em abstract pattern matching} in COSMOS/Scope to 
an industrial benchmark where we obtained up to 3 orders of magnitude reduction 
in shuffled data and 1.23x average speedup in total processing time.
% and 1.09x average speedup in end-to-end latency
\end{abstract}

\input{01_intro}
\input{02_related_work}
\input{03_design}
\input{04_implementation}
\input{05_evaluation}
\input{06_conclusions}


\small
\bibliographystyle{abbrv}
\bibliography{mix}




\end{document}
